---
title: "Create LAS Script"
author: "Mick Morrison"
date: "2025-06-29"
format:                # Quarto uses `format`, not `output`
  html:
    self-contained: true
    theme: spacelab
server: shiny          # activates a Shiny back-end
execute:
  echo: true
  message: false
  warning: false
editor: visual
bibliography: references.bib
---

```{r setup, include=FALSE}
# make rgl render to WebGL widgets rather than an on-screen device

options(rgl.useNULL = TRUE)

rgl::setupKnitr(autoprint = TRUE)
```

# About

This script creates a synthetic LAS dataset with an assigned Coordinate Reference System (CRS) for development and testing purposes. This includes what I hope is a compliant header. It draws on *R* [@base], *lidR* [@roussel2020], *sf* [@sf] and *rgl* [@rgl] and was created with help from ChatGPT o3. A simple visualisation of the generated dataset is provided.

Copy and run the script below to use. The test las can be found in this repo at \[`/synth_las.las`\].

```{r, include=TRUE}

#required packages
library(lidR)
library(sf)
library(rgl)

## 1  generate synthetic cloud
set.seed(1)
n_pts <- 100000 ## Set the number of points. 1000000 defaul
df <- data.frame(
  X = runif(n_pts, 0, 100),
  Y = runif(n_pts, 0, 100),
  Z = runif(n_pts, 0, 30),
  Intensity       = sample(1:255, n_pts, TRUE),
  ReturnNumber    = 1L,
  NumberOfReturns = 1L,
  Classification  = 1L
)

## 2  build LAS & patch header
las <- LAS(df)
las@header@PHB$`File Source ID` <- 1L
las@header@PHB$`Version Major`  <- 1L
las@header@PHB$`Version Minor`  <- 2L
projection(las) <- 28356               # GDA94 / MGA 56
writeLAS(las, "synth_las.las", index = TRUE)
```

## View the test LAS

```{r}
## 3  save & display (widget auto-prints)
plot3d(las$X, las$Y, las$Z, col = las$Intensity, size = 3)
#rglwidget()
```

## Download the test LAS

[/synth_las.las](/synth_las.las)
